---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.path   = "man/figures/README-",
  fig.width  = 8,
  fig.height = 1.2,
  out.width  = "100%"
)
```

# ggspruce

<!-- badges: start -->
<!-- badges: end -->

ggspruce provides functions that make automatic adjustments to ggplot2 plots
to improve plot aesthetics.

## Installation

You can install the development version of ggspruce from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("sheridar/ggspruce")
```

<br>

## Optimizing color palettes

This is a color palette from the MetBrewer package.
This palette can be improved by modifying colors 2, 12, and 13 to make
them more distinct.

```{r}
library(ggspruce)
library(ggplot2)
library(MetBrewer)
library(cowplot)

clrs <- as.character(met.brewer("Austria", 15))

plot_colors(clrs)
```

The `spruce_up_colors()` function calculates the pairwise distance between
colors in the palette, and identifies pairs of colors that are very similar.
The difference threshold used to identify similar colors can be modified with
the `difference` parameter, with higher values resulting in more distinct colors.
The color palette is optimized by modifying the color attribute specified
with the `adjust` parameter.

In the example below, color lightness is automatically adjusted for sets of
colors that do not meet the difference threshold.

```{r}
new_clrs <- clrs |>
  spruce_up_colors(
    difference = 10,
    adjust     = "lightness"
  )

plot_colors(new_clrs)
```

<br>

The colors can be made more distinct by modifying the hue instead of lightness
and/or increasing the difference threshold.

```{r}
new_clrs <- clrs |>
  spruce_up_colors(
    difference = 15,
    adjust     = "hue"
  )

plot_colors(new_clrs)
```

<br>

Color palettes can be automatically modified on the fly using the
`scale_color_spruce()` and `scale_fill_spruce()` functions.
Original colors are shown on the left and adjusted colors on the right.

```{r, fig.width = 8, fig.height = 5}
clrs <- as.character(met.brewer("Nattier", 8))

# Original colors
plt <- diamonds |>
  ggplot(aes(carat, price, color = clarity)) +
  geom_point(size = 1) +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.position = "bottom"
  )

plt1 <- plt +
  scale_color_manual(values = clrs) +
  ggtitle("original")

# Automatically adjust colors
plt2 <- plt +
  scale_color_spruce(
    values     = clrs,
    difference = 25,
    adjust     = "lightness"
  ) +
  ggtitle("adjusted")

plot_grid(plt1, plt2, nrow = 1)
```

<br>

## Colorblind filters

Color similarity can be assessed using colorblind filter(s) and colors adjusted
to increase perceived differences.

```{r, fig.width = 8, fig.height = 6}
library(colorspace)

clrs <- as.character(met.brewer("Moreau", 10))

new_clrs <- clrs |>
  spruce_up_colors(
    difference = 16,
    adjust     = "lightness",
    filter     = "deutan"
  )

# Plot adjusted colors without filter
plt1 <- clrs |>
  plot_colors() +
  ggtitle("original")

plt2 <- new_clrs |>
  plot_colors() +
  ggtitle("adjusted")

# Plot adjusted colors with colorblind filter
plt3 <- clrs |>
  deutan() |>
  plot_colors() +
  ggtitle("original + filter")
  
plt4 <- new_clrs |>
  deutan() |>
  plot_colors() +
  ggtitle("adjusted + filter")
  
plot_grid(plt1, plt2, plt3, plt4, ncol = 1)
```

